---
title: "Advanced Methods for Regression and Classification - Exercise 6 Experiment - Dzenan Hamzic, TU Wien"
output: html_notebook
---

## Bone Mineral Density Data

### Description
Measurements in the bone mineral density of 261 north american adolescents, as function of age. Each value is the difference in spnbmd taken on two consecutive visits, divided by the average. The age is the average age over the two visits.

```{r}
#install.packages("ElemStatLearn")
library(splines)
set.seed(123)
data(bone, package="ElemStatLearn")
bone
```


```{r}
summary(bone)
```

#### make binary variables from categorical
```{r}

#data <- dummy_cols(bone, select_columns = "gender")
data <- bone[which(bone$gender=='male'),]
#detach(bone)
#data$gender<-NULL
head(data)
```



#### Split data randomly with 70/30 splits (train and test)
```{r}
n <- dim(data)[1]
is.train <- sample(1:n, ceiling(2./3. * n))
data.train <- data[is.train,]
data.test <- data[-is.train,]

data.train
```

## 1. Polynomial Regression

## 1.a)
```{r}
lmod2 = loess(spnbmd ~ age, data=data.train, span = 0.3)
lmod5 = loess(spnbmd ~ age, data=data.train, span = 0.6)
lmod7 = loess(spnbmd ~ age, data=data.train, span = 0.9)
```


```{r}
# get smoothed output
smoothed2 <- predict(lmod2) 
smoothed5 <- predict(lmod5) 
smoothed7 <- predict(lmod7) 
```

#### plot
```{r}
#plot(bone$spnbmd, x=bone$age, type="l", main="Loess Smoothing and Prediction", xlab="Date", ylab="Unemployment (Median)")
scatter.smooth(data$spnbmd ~ data$age, span = 0.3, col="red", main="Loess Smoothing 0.2 and Prediction")
scatter.smooth(data$spnbmd ~ data$age, span = 0.6, main="Loess Smoothing 0.5 and Prediction")
scatter.smooth(data$spnbmd ~ data$age, span = 0.9, main="Loess Smoothing 0.7 and Prediction")
```



#### predict
```{r}
test_pred_lmod2 <- predict(lmod2, newdata = data.test)
test_pred_lmod5 <- predict(lmod5, newdata = data.test)
test_pred_lmod7 <- predict(lmod7, newdata = data.test)
```



## 1.b)
#### 0.3 span
```{r}
plot(data.test$age, test_pred_lmod2, xlab='age', ylab='spnbmd', xlim=range(data.test$age), ylim=range(data.test$spnbmd),col="red")
par(new=T)
scatter.smooth(data.test$spnbmd ~ data.test$age, xlab='', ylab='', span = 0.3, main="Predicted (Red) vs Actual (Black) - Span 0.3", xlim=range(data.test$age), ylim=range(data.test$spnbmd))

### MSE
mean((test_pred_lmod2 - data.test$spnbmd)^2, na.rm = TRUE)
```

#### 0.6 span
```{r}
plot(data.test$age, test_pred_lmod5, xlab='age', ylab='spnbmd', xlim=range(data.test$age), ylim=range(data.test$spnbmd),col="red")
par(new=T)
scatter.smooth(data.test$spnbmd ~ data.test$age, xlab='', ylab='', span = 0.6, main="Predicted (Red) vs Actual (Black) - Span 0.6", xlim=range(data.test$age), ylim=range(data.test$spnbmd))

### MSE
mean((test_pred_lmod5 - data.test$spnbmd)^2, na.rm = TRUE)
```
#### 0.9 span
```{r}
plot(data.test$age, test_pred_lmod7, xlab='age', ylab='spnbmd', xlim=range(data.test$age), ylim=range(data.test$spnbmd),col="red")
par(new=T)
scatter.smooth(data.test$spnbmd ~ data.test$age, xlab='', ylab='', span = 0.9, main="Predicted (Red) vs Actual (Black) - Span 0.9", xlim=range(data.test$age), ylim=range(data.test$spnbmd))

### MSE
mean((test_pred_lmod7 - data.test$spnbmd)^2, na.rm = TRUE)
```


## 2. B-splines
```{r}
bs3 <- bs(data.test$age, df=3)
bs6 <- bs(data.test$age, df=9)
bsCV <- smooth.spline(data.test$age, data.test$spnbmd)

matplot(bs3, type="l", lty=1)
matplot(bs6, type="l", lty=1)
```
### 2.a)
```{r}
lm2 = lm(data.test$spnbmd ~ bs3)
lm4 = lm(data.test$spnbmd ~ bs6)

x <- sort(data.test$age)
y1 <- lm2$fitted.values[order(data.test$age)]
y2 <- lm4$fitted.values[order(data.test$age)]

# lines function produces ERROR in my setting

scatter.smooth(x,y1, main="LS Regression with B-splines - basis functions:3")
scatter.smooth(x,y2, main="LS Regression with B-splines - basis functions:9")
```
### 2.b)
```{r}
spnbmd_predicted_bs2 = predict.lm(lm2, list(x=data.test$age))
spnbmd_predicted_bs4 = predict.lm(lm4, list(x=data.test$age))
spnbmd_predicted_bsCV = predict(bsCV, x=data.test$age)

plot(data.test$age, data.test$spnbmd, xlab='age', ylab='spnbmd', xlim=range(data.test$age), ylim=range(data.test$spnbmd),col="red", main="LS Regression with B-splines - Prediction vs. Actual")
par(new=T)
scatter.smooth(spnbmd_predicted_bs2 ~ data.test$age, col="blue", xlab='', ylab='', xlim=range(data.test$age), ylim=range(data.test$spnbmd))


plot(data.test$age, data.test$spnbmd, xlab='age', ylab='spnbmd', xlim=range(data.test$age), ylim=range(data.test$spnbmd),col="red", main="LS Regression with B-splines - Prediction vs. Actual")
par(new=T)
scatter.smooth(spnbmd_predicted_bs4 ~ data.test$age, col="blue", xlab='', ylab='', xlim=range(data.test$age), ylim=range(data.test$spnbmd))


plot(data.test$age, data.test$spnbmd, xlab='age', ylab='spnbmd', xlim=range(data.test$age), ylim=range(data.test$spnbmd),col="red", main="LS Regression with CV B-splines - Prediction vs. Actual")
par(new=D)
scatter.smooth(spnbmd_predicted_bsCV, xlab='age', ylab='spnbmd', xlim=range(data.test$age), ylim=range(data.test$spnbmd), main='')
```


```{r}
### MSE
mean((spnbmd_predicted_bs2 - data.test$spnbmd)^2, na.rm = TRUE)
mean((spnbmd_predicted_bs4 - data.test$spnbmd)^2, na.rm = TRUE)
mean((spnbmd_predicted_bsCV$y - data.test$spnbmd)^2, na.rm = TRUE)
```

## 3. Natural Cubic Splines
```{r}
ns1 <- ns(data.test$age, df=3)
ns2 <- ns(data.test$age, df=9)

matplot(ns1, type="l", lty=1)
matplot(ns2, type="l", lty=1)
```
### 3.a)
```{r}
lm5 = lm(data.test$spnbmd ~ ns1)
lm6 = lm(data.test$spnbmd ~ ns2)

x <- sort(data.test$age)
y1 <- lm5$fitted.values[order(data.test$age)]
y2 <- lm6$fitted.values[order(data.test$age)]

# lines function produces ERROR in my setting

scatter.smooth(x,y1, main="Natural cubic splines - fitted 3 splines")
scatter.smooth(x,y2, main="Natural cubic splines - fitted 9 splines")
```
### 2.b)
```{r}
spnbmd_predicted_ns1 = predict.lm(lm5, list(x=data.test$age))
spnbmd_predicted_ns2 = predict.lm(lm6, list(x=data.test$age))

plot(data.test$age, data.test$spnbmd, xlab='age', ylab='spnbmd', xlim=range(data.test$age), ylim=range(data.test$spnbmd),col="red", main="LS Regression with (3) Natural-splines - Prediction vs. Actual")
par(new=T)
scatter.smooth(spnbmd_predicted_ns1 ~ data.test$age, col="blue", xlab='', ylab='', xlim=range(data.test$age), ylim=range(data.test$spnbmd))


plot(data.test$age, data.test$spnbmd, xlab='age', ylab='spnbmd', xlim=range(data.test$age), ylim=range(data.test$spnbmd),col="red", main="LS Regression with (9) Natural-splines - Prediction vs. Actual")
par(new=T)
scatter.smooth(spnbmd_predicted_ns2 ~ data.test$age, col="blue", xlab='', ylab='', xlim=range(data.test$age), ylim=range(data.test$spnbmd))
```

```{r}
### MSE
mean((spnbmd_predicted_ns1 - data.test$spnbmd)^2, na.rm = TRUE)
mean((spnbmd_predicted_ns2 - data.test$spnbmd)^2, na.rm = TRUE)
```